Body size patterns

### 2020-09-22 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
first ms draft

---
title: "Body size patterns"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

@@

```{r Load data, message=FALSE}
library(itsadug)
library(arm)
library(relaimpo)
library(GGally)
library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

@@

```{r gams for body size - data prep}
speciesIn <- 'bnt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the speces and river smooths

@@

@@

@@

@@

```{r length ageInSamples = 1}

@@

#BNT
lenModBNT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bnt', ageInSamples == 1))
lenModBNT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT3 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT4 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday * riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT5 <- lm(observedLength ~ 0 + factor(cohort) * riverGG * yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))

@@

@@

cs_ais1 <- left_join(cohortStrength,ais1)

@@

ggplot(cs_ais1 %>% filter(meanYday > 0, species %in% c('bkt','bnt')), aes(cohortStrengthScaled, meanLen)) +
  geom_point(aes(size = meanYday)) +
  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color = 'Cohort') +
  scale_y_continuous("Mean Fork Length (mm)") +
  scale_x_continuous("Cohort Strength") +
  facet_grid(speciesGG ~ metaPop01)

@@

@@

```{r models with season-don't run. gam structure given best model from above}

>>>>

---
title: "Body size patterns"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

@@

```{r Load data, message=FALSE}
library(itsadug)
library(gratia)
library(arm)
library(relaimpo)
library(GGally)
# library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

@@

```{r theme_publication}

theme_publication <- function(base_size=14, base_family="") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(),
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.2, "cm"),
            #    legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold"),
            panel.border = element_rect(colour = "black")
    ))

}
```

```{r gams for body size - data prep}
speciesIn <- 'bkt'
depVarIn <- 'observedLength'
kIn <- 4 # # of knots for the species and river smooths

@@

########################################
# tags that may need cohort fixes

#           tag
# 1  00088cc1c3
# 2  00088cc5c8
# 3  00088d0a10
# 4  00088d1a28
# 5  00088d20f5
# 6  00088d215f
# 7  00088d23b4
# 8  0009f6ecc2
# 9  0009f6edf1
# 10 0009f6ee11
# 11 0009f6f33c
# 12 0009f6f404
# 13 1bf20e52f0
# 14 1bf20ebfa7
########################################
```

@@

# temporary fix - added 9/18/19, use this just to remake AIC1... fig
#vB[vB$tag == '0009f6f33c',]$cohort <- 2014  
#vB[vB$tag == '0009f6f404',]$cohort <- 2014

@@

```{r raw data tables}

# get proportion of fish tagged for the first time as age0
dataAllSpp = vB %>% filter(cohort %in% 2000:2013, enc == 1) %>%
  filter(!(cohort < 2002 & riverFactor %in% c('wb jimmy','wb mitchell','wb obear')))

AIS <- dataAllSpp %>%
  group_by(tag) %>%
  summarise(minAIS = min(ageInSamples))

table(AIS$minAIS)/nrow(AIS)

# counts by season, year, spp
counts <- dataAllSpp %>%
  group_by(season, year, species) %>%
  summarize(n = n())

@@

```{r length ageInSamples equals one}

@@

################################################
# compare mean lengths between trout
meansWide <- meanCVLen %>%
  select(-sdL) %>%
  pivot_wider(names_from = species, values_from = meanL) %>%
  filter(riverGG %in% c('West Brook', 'Open Large'))

ggplot(meansWide, aes(bkt, bnt)) +
  geom_point() +
  theme_publication() +
  scale_x_continuous('Brook Trout mean length (mm)') +
  scale_y_continuous('Brown Trout mean length (mm)', limits = c(60,90)) +
  geom_smooth(method = 'lm', color = 'black') +
  facet_wrap(~riverGG)

ggsave('./figures/AIS1_MeanLengthBKT_BNT.png', width = 11, height = 6, units='in')

#summary(lm(bnt ~ bkt, data = filter(meansWide, riverGG == 'West Brook')))
#summary(lm(bnt ~ bkt, data = filter(meansWide, riverGG == 'Open Large', bnt < 90)))
corr(meansWide %>% ungroup() %>% filter(riverGG == 'West Brook') %>% select(bkt, bnt) %>% drop_na())
corr(meansWide %>% ungroup() %>% filter(riverGG == 'Open Large', bnt < 90) %>% select(bkt, bnt) %>% drop_na())
#########################
#BNT
lenModBNT1 <- lm(observedLength ~  factor(cohort) * factor(riverGG), data = vB %>% filter(cohort %in% 2000:2013, species == 'bnt', ageInSamples == 1))
lenModBNT2 <- lm(observedLength ~  factor(cohort) + riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT3 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT4 <- lm(observedLength ~ 0 + factor(cohort) * riverGG + yday * riverGG, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))
# lenModBNT5 <- lm(observedLength ~ 0 + factor(cohort) * riverGG * yday, data = vB %>% filter(cohort %in% 2000:2013,  species == 'bnt', ageInSamples == 1))

@@

```{r corr btw bkt/bnt mean lengths all samples}
meanCVLenAll <- 
#vB %>% filter(cohort %in% 2000:2013) %>% # leave out 2013 cohort - some bad cohort assignments I've asked Matt to check 8/4/2020
  vB %>% filter(cohort %in% 2000:2012) %>%
  group_by(species, riverGG, cohort, ageInSamples) %>%
  summarize(meanL = mean(observedLength),
            sdL = sd(observedLength))

################################################
# compare mean lengths between trout
meansWideAll <- meanCVLenAll %>%
  select(-sdL) %>%
  pivot_wider(names_from = species, values_from = meanL) %>%
  filter(riverGG %in% c('West Brook', 'Open Large'))

# get correlations
corrMod <- function(df) {
  df <- df %>% select(bkt, bnt) %>% drop_na()
  corr(df)
}

byAIS_River <- meansWideAll %>%
  group_by(riverGG, ageInSamples) %>%
  nest()

byAIS_River <- byAIS_River %>%
  mutate(corr = map(data, corrMod)) %>%
  unnest() %>%
  distinct(riverGG, ageInSamples, corr)

meansWideAll$AIS_labels <- factor(meansWideAll$ageInSamples, labels = c('Autumn, 0', 'Winter, 0', 'Spring, 1', 'Summer, 1', 'Autumn, 1', 'Winter, 1', 'Spring, 2', 'Summer, 2', 'Autumn, 2'))

byAIS_River$AIS_labels <- factor(byAIS_River$ageInSamples, labels = c('Autumn, 0', 'Winter, 0', 'Spring, 1', 'Summer, 1', 'Autumn, 1', 'Winter, 1', 'Spring, 2', 'Summer, 2', 'Autumn, 2'))

ggplot(meansWideAll, aes(bkt, bnt)) +
  geom_point() +
  theme_publication() +
  scale_x_continuous('Brook Trout mean length (mm)') +
  scale_y_continuous('Brown Trout mean length (mm)') +
  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
  geom_abline(intercept = 0, slope = 1, color = 'grey') +
  geom_text(aes(label = round(corr, digits = 2), x = 95, y = 235), size = 3.5, data = byAIS_River) +
  facet_grid(riverGG ~ AIS_labels)

ggsave('./figures/AISALL_MeanLengthBKT_BNT.png', width = 11, height = 6, units='in')

@@

cohortStrength2 <- cohortStrength
cohortStrength2$speciesGG <- factor(cohortStrength2$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

cs_ais1 <- left_join(cohortStrength2,ais1)

@@

cs_ais1$metaPopLabel <- factor(cs_ais1$metaPop01, label = c("Isolated Large", "MetaPopulation"))

ggplot(cs_ais1 %>% filter(meanYday > 0, species %in% c('bkt','bnt')), aes(cohortStrengthScaled, meanLen)) +
  geom_point(aes(size = meanYday)) +
  geom_smooth(method = 'lm', se = FALSE, color = 'black') +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        #strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color = 'Cohort') +
  scale_y_continuous("Mean Fork Length (mm)") +
  scale_x_continuous("Cohort Strength") +
  facet_grid(speciesGG ~ metaPopLabel)

@@

appraise(modBKT)
ggsave('./figures/appraiseBKT.png', width = 11, height = 8, units='in')

appraise(modBNT)
ggsave('./figures/appraiseBNT.png', width = 11, height = 8, units='in')

@@

```{r models with season-do not run. gam structure given best model from above}
